Arrows for Parallel Computations

### 2017-09-25 paper/content/motivation.tex
wip restructuring the paper, wip on motivation's contribution

\paragraph{Contributions}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose a new Arrows-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]|, which converts a list of Arrows into a new parallel Arrow. Because of this, we do not lose any benefits of using Arrows as the parallelism is encapsulated as another combinator instead of a different type. The resulting Arrow can be used in the same way as a potential serial version.

This is a big advantage as we do not introduce any new types like Monad solutions like the |Par| Monad do. We can just \enquote{plug} in parallel parts into sequential Arrow-based programs without having to change anything.

We host this functionality in the |ArrowParallel| typeclass, which abstracts all parallel implementation logic in a way that the backend can be swapped easily. This means it is possible to write code against the interface of a common typeclass without being bound to any specific backend. So as an example, during development, we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance.

\ptcomment{* We evaluate the expressive power of  arrow parallelism. Showing that 
  + there is no need to introduce additional types
  + that it is possible to define algorthmic skeletons (Section 6,7)
  + ...

* We demonstrate that Arrow Parallelism can exploit multiple parallel Haskell implementations, and with low overhead, i.e. less than X%. 
... (Section 8) }

>>>>

\paragraph{Contributions}
%
%\olcomment{HIT HERE REALLY STRONG}
%
%\subsection{Impact of parallel Arrows}
%\olcomment{move this to Contributions in the front or something}
We propose a new Arrow-based encoding for parallelism based on a new Arrow combinator |parEvalN :: [arr a b] -> arr [a] [b]|, which converts a list of Arrows into a new parallel Arrow. Because of this, we do not lose any benefits of using Arrows as the parallelism is encapsulated as another combinator instead of a different type. The resulting Arrow can be used in the same way as a potential serial version.

This is a big advantage as we do not introduce any new types as Monad solutions like the |Par| Monad do. We can just \enquote{plug} in parallel parts into sequential Arrow-based programs without having to change integral parts of the code.

We host this functionality in the |ArrowParallel| typeclass, which abstracts all parallel implementation logic in a way that the backend easily be swapped. This means it is possible to write code against the interface of a common typeclass without being bound to any specific backend. So as an example, during development, we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance.

We evaluate the expressive power of arrow parallelism further by showing that it is possible to define algorithmic skeletons with Arrows (Sections \ref{sec:map-skeletons},~\ref{sec:topology-skeletons}).

Finally, we demonstrate that Arrow Parallelism is a viable alternative to existing approaches and prove that only low overhead is introduced, i.e. less than x\% \mbcomment{fix this after benchmarks are finally done} in Section \ref{sec:benchmarks}.

%\ptcomment{* We evaluate the expressive power of  arrow parallelism. Showing that 
%  + there is no need to introduce additional types
%  + that it is possible to define algorthmic skeletons (Section 6,7)
%  + ...

%* We demonstrate that Arrow Parallelism can exploit multiple parallel %Haskell implementations, and with low overhead, i.e. less than X%. 
%... (Section 8) }

### 2017-09-25 paper/content/parallelHaskells.tex
wip restructuring the paper, wip on motivation's contribution

\begin{flushright}

\end{flushright}\section{Background}
\label{sec:background}
\subsection{Short introduction to parallel Haskells}
\label{sec:parEvalNIntro}
There are already several ways to write parallel programs in Haskell. As we base our parallel arrows on existing parallel Haskells, we will now give a short introduction to the ones we use as backends in this paper.

In its purest form, parallel computation (on functions) can be looked at as the execution of some functions |a -> b| in parallel or |parEvalN :: [a -> b] -> [a] -> [b]|, as also Figure~\ref{fig:parEvalN} symbolically shows.
\begin{figure}[t]
  \centering
	\includegraphics[scale=0.7]{images/parEvalN}
	\caption{Schematic illustration of |parEvalN|.}
	\label{fig:parEvalN}
\end{figure}
%Before we go into detail on how we can use this idea of parallelism
%for parallel Arrows, as a short introduction to parallelism in
%Haskell we will now implement |parEvalN| with several different
%parallel Haskells.
As a demonstration, we implement here the non-Arrows |parEvalN| in multiple parallel Haskells.

>>>>

\subsection{Short introduction to parallel Haskells}
\label{sec:parEvalNIntro}
There are already several ways to write parallel programs in Haskell. As we base our parallel Arrows on existing parallel Haskells, we will now give a short introduction to the ones we use as backends in this paper.

In its purest form, parallel computation (on functions) can be looked at as the execution of some functions |a -> b| in parallel or |parEvalN :: [a -> b] -> [a] -> [b]|, as also Figure~\ref{fig:parEvalN} symbolically shows.
\begin{figure}[t]
  \centering
	\includegraphics[scale=0.7]{images/parEvalN}
	\caption{Schematic illustration of |parEvalN|.}
	\label{fig:parEvalN}
\end{figure}
%Before we go into detail on how we can use this idea of parallelism
%for parallel Arrows, as a short introduction to parallelism in
%Haskell we will now implement |parEvalN| with several different
%parallel Haskells.
We implement this non-Arrow version here which will later in this paper be adapted for usage in our Arrow based parallel Haskell. Furthermore, the implementation of |parEvalN| serves as a short demonstration of how the used parallel Haskell backends work.
